---
title: "Interpretable Analysis of School Policy Decisions, Comparison to schools outside the program"
author: "Charles Saluski"
# date: "1/4/2022"
output: pdf_document
---

```{r}
library(data.table)
library(dplyr)
library(stringr)
library(purrr)
library(openxlsx)
```

```{r}
district.map.dt <- fread("Data Sources CSV/building_map_data/District MAP content area and grade all disag.csv")

district.map.dt[, State.District.ID := paste0("MO-", str_pad(COUNTY_DISTRICT, 6, "left", "0"))]

dci.buildings.dt <- fread("Data Sources/DCI Data/Active Districts/Active_DCI_buildings_2017_2022.csv")

dci.districts.dt <- unique(dci.buildings.dt[, .(State.District.ID, currentSchoolYear)])

dci.districts.dt <- dci.districts.dt[order(currentSchoolYear)]

# Account for years taken off from the program by counting along the years the
# district was actually in the program
dci.districts.dt[
  ,
  year.in.program := which(.SD$currentSchoolYear == currentSchoolYear),
  by = "State.District.ID"]

dci.years.dt <- dci.districts.dt

dci.districts.dt <- dci.districts.dt[district.map.dt, on=c("State.District.ID", "currentSchoolYear" = "YEAR")]

dci.districts.dt[, year.in.program := ifelse(is.na(year.in.program), 0, year.in.program)]

dci.districts.dt <- dci.districts.dt[
  (currentSchoolYear == 2019 | currentSchoolYear == 2021)
  & TYPE %in% c("Total", "Non IEP Students", "IEP Non MAPA", "IEP_student")]

dci.districts.dt <- dci.districts.dt[CONTENT_AREA == "Eng. Language Arts" & GRADE_LEVEL == "03"]

year.in.program.dt <- unique(dci.districts.dt[, .(State.District.ID, currentSchoolYear, year.in.program, COUNTY_DISTRICT)])

state.average.dt <- fread("./Data Sources CSV/MAP State Average.csv")
state.average.dt[Type == "IEP (exclude MAPA)", Type := "IEP Non MAPA"]
# state.average.dt[Type == "Total (exclude MAPA)", Type := "IEP Non MAPA"]

state.average.dt[, PROF_AND_ADV_PCT_STATE := ((Advanced + Proficient) / Reportable) * 100]
state.average.dt <- state.average.dt[
  Year >= 2019 & Type %in% unique(dci.districts.dt$TYPE),
  .(Year, Type, PROF_AND_ADV_PCT_STATE)]
```

```{r}
discipline.dt <- fread("Data Sources CSV/District Discipline.csv")
attendance.dt <- fread("Data Sources CSV/District Proportional Attendance Rates.csv")

discipline.attendance.dt <- discipline.dt[attendance.dt, on=c("YEAR", "COUNTY_DISTRICT_CODE", "DISTRICT_NAME")]
discipline.attendance.dt <- discipline.attendance.dt[(YEAR == 2019 | YEAR == 2021) & complete.cases(discipline.attendance.dt)]

discipline.attendance.dt <- discipline.attendance.dt[, .(YEAR, COUNTY_DISTRICT_CODE, DSCPLN_INCIDENT_RATE, PROPORTIONAL_ATTENDANCE_K_8_PCT, PROPORTIONAL_ATTENDANCE_IEP_PCT)]

for (col in colnames(discipline.attendance.dt)) {
  if (grepl("PROPORTIONAL", col)) {
    num.col <- as.numeric(discipline.attendance.dt[, get(col)])
    num.col <- ifelse(is.na(num.col), 0, num.col)
    discipline.attendance.dt[, paste(col) := num.col]
  }
}

discipline.map.dt <- dci.districts.dt[discipline.attendance.dt, on=c("currentSchoolYear" = "YEAR", "COUNTY_DISTRICT" = "COUNTY_DISTRICT_CODE")]

# TODO merge discipline with DCI Districts DT
discipline.map.dt <- discipline.map.dt[complete.cases(discipline.map.dt)]


# discipline.diff.dt <- merge(discipline.map.dt, discipline.map.dt,
#   by = c("State.District.ID"),  
#   suffixes = c("_2019", "_2021")
#   )[currentSchoolYear_2019 == 2019 & currentSchoolYear_2021 == 2021,
#   ]


# diff.discipline.cols <- names(discipline.map.dt)[as.vector(sapply(discipline.map.dt, class)) %in% c("numeric", "num", "int")]

# for (col in diff.discipline.cols) {
#   diff.col.name <- paste(col, "diff", sep=".")
#   col.x.name <- c(paste(col, "2019", sep="_"))
#   col.y.name <- c(paste(col, "2021", sep="_"))
#   # I can't get this to work just using the variable names, so we're using get
#   discipline.diff.dt[[diff.col.name]] <- discipline.diff.dt[, get(col.y.name) - get(col.x.name)]
# }

```

```{r}

dci.districts.dt <- dci.districts.dt[state.average.dt, on = c("currentSchoolYear" = "Year", "TYPE" = "Type")]

for (col in names(dci.districts.dt)) {
  if (grepl("PCT", col)) {
    dci.districts.dt[, paste0(col) := as.numeric(get(col))]
  }
}


dci.districts.dt[, PROF_AND_ADV_PCT := PROFICIENT_PCT + ADVANCED_PCT]

# get changes from 2019 to 2021
covid.diff.dt <- merge(dci.districts.dt, dci.districts.dt,
  by = c("State.District.ID", "TYPE", "GRADE_LEVEL", "CONTENT_AREA", "SCHOOL_NAME", "SCHOOL_CODE", "DISTRICT_NAME", "COUNTY_DISTRICT", "CATEGORY", "SUMMARY_LEVEL"),  
  suffixes = c("_2019", "_2021")
  )[currentSchoolYear_2019 == 2019 & currentSchoolYear_2021 == 2021,
  ]

diff.cols <- names(dci.districts.dt)[as.vector(sapply(dci.districts.dt, class)) %in% c("numeric", "num", "int")]

for (col in diff.cols) {
  diff.col.name <- paste(col, "diff", sep=".")
  col.x.name <- c(paste(col, "2019", sep="_"))
  col.y.name <- c(paste(col, "2021", sep="_"))
  # I can't get this to work just using the variable names, so we're using get
  covid.diff.dt[[diff.col.name]] <- covid.diff.dt[, get(col.y.name) - get(col.x.name)]
}

covid.diff.dt[, gap.2019 := PROF_AND_ADV_PCT_2019 - PROF_AND_ADV_PCT_STATE_2019]
covid.diff.dt[, gap.2021 := PROF_AND_ADV_PCT_2021 - PROF_AND_ADV_PCT_STATE_2021] 
covid.diff.dt[, gap.diff := gap.2021 - gap.2019]
```
See how much the proficient and advanced percent changed before and after covid,
and see if its change was impacted by the number of years in the program.
```{r}
year.list <- list()
box.list <- list()

diff.types <- c("Total", "Non IEP Students", "IEP_student")

for (year in unique(covid.diff.dt$year.in.program_2019)) {
  for (type in unique(covid.diff.dt$TYPE))
  {
    year.list[[paste(type, year)]] <- data.table(
      value = mean(covid.diff.dt[year.in.program_2019 == year & TYPE == type]$PROF_AND_ADV_PCT.diff),
      year,
      type
      )
    box.list[[paste(type, year)]] <- data.table(covid.diff.dt[year.in.program_2019 == year & TYPE == type, .(PROF_AND_ADV_PCT.diff, year, type, PROF_AND_ADV_PCT_STATE.diff, State.District.ID, gap.diff)])
  }
}

# diff.interest.cols <- c("PROPORTIONAL_ATTENDANCE_TOTAL_PCT", "PROPORTIONAL_ATTENDANCE_IEP_PCT", "DSCPLN_INCIDENT_RATE")

# for (year in unique(covid.diff.dt$year.in.program_2019)) {
#   for (col in diff.interest.cols)
#   {
#     year.list[[paste(col, year)]] <- data.table(
#       value = mean(covid.diff.dt[year.in.program_2019 >= year, get(paste0(col, ".diff"))]),
#       year,
#       type = col
#       )
#   }
# }

covid.diff.res.dt <- data.table(do.call(rbind, year.list))
covid.diff.res.dt$year <- as.factor(covid.diff.res.dt$year)
covid.box.dt <- data.table(do.call(rbind, box.list))
covid.box.dt[, n := paste(year, ", n = ", length(unique(.SD$State.District.ID)), sep = ""), by = year]
covid.box.dt$year <- as.factor(covid.box.dt$year)
```


```{r}
library(ggplot2)

dest <- "./img_out/by_iep_groups/"
if (!dir.exists(dest)) {
  dir.create(dest)
}

map.change.plot <- ggplot() +
  geom_boxplot(
    data = covid.box.dt,
    mapping = aes(x = n, y = PROF_AND_ADV_PCT.diff, fill = type)
    ) +
  geom_errorbar(
    data = unique(covid.box.dt[, .(n, type, PROF_AND_ADV_PCT_STATE.diff)]),
    mapping = aes(
      x = n,
      ymin = PROF_AND_ADV_PCT_STATE.diff, ymax = PROF_AND_ADV_PCT_STATE.diff,
      color = type)
    ) +
  xlab("Years in Program in 2019, n = number of districts") +
  ylab("3rd Grade MAP ELA difference, 2019 to 2021") +
  labs(title = "MAP ELA changes by years in DCI program", fill = "MAP Group", color = "State Average") 


filename <- paste(dest, "map_ela_changes.png", sep = "")
png(filename = filename, width = 6, height = 8, unit = "in", res = 200)
plot(map.change.plot)
dev.off()
```


```{r}
gap.diff.plot <- ggplot(data = covid.box.dt) +
  geom_boxplot(aes(x=n, y=gap.diff, fill=type)) +
  xlab("Years in Program in 2019, n = number of districts") +
  ylab("3rd Grade MAP ELA State Average Difference, 2019 to 2021") +
  labs(title = "MAP ELA gap changes by years in DCI program, \n normalized against state average", fill = "MAP Group", color = "State Average") 


filename <- paste(dest, "map_ela_diff_changes.png", sep = "")
png(filename = filename, width = 6, height = 8, unit = "in", res = 200)
plot(gap.diff.plot)
dev.off()
```

```{r}

dci.districts.dt[, n := paste(year.in.program, ", n = ", length(unique(.SD$State.District.ID)), sep = ""), by = c("year.in.program")]

map.years.plot <- ggplot() +
  geom_boxplot(
    data = dci.districts.dt,
    mapping = aes(x = n, y = PROF_AND_ADV_PCT, fill = TYPE)
    ) +
  geom_errorbar(
    data = unique(dci.districts.dt[, .(n, TYPE, PROF_AND_ADV_PCT_STATE, currentSchoolYear)]),
    mapping = aes(
      x = n,
      ymin = PROF_AND_ADV_PCT_STATE, ymax = PROF_AND_ADV_PCT_STATE,
      color = TYPE)
    ) +
  facet_grid(currentSchoolYear ~ .) +
  xlab("Years in Program in year, n = number of districts in 2021") +
  ylab("3rd Grade MAP ELA score") +
  labs(title = "MAP ELA boxplot by years in DCI program in 2021", fill = "MAP Group", color = "State Average") 

filename <- paste(dest, "map_ela_by_year.png", sep = "")
png(filename = filename, width = 6, height = 8, unit = "in", res = 200)
plot(map.years.plot)
dev.off()
```

```{r}

discipline.map.dt[, n := paste(year.in.program, ", n = ", length(unique(.SD$State.District.ID)), sep = ""), by = c("year.in.program")]


discipline.years.plot <- ggplot() +
  geom_boxplot(
    data = discipline.map.dt,
    mapping = aes(x = n, y = DSCPLN_INCIDENT_RATE)
    ) +
  facet_grid(currentSchoolYear ~ .) +
  xlab("Years in Program in year, n = number of districts in 2021") +
  ylab("") +
  labs(title = "Discipline rate boxplot by years in DCI program")

filename <- paste(dest, "discipline_by_years.png", sep = "")
png(filename = filename, width = 6, height = 8, unit = "in", res = 200)
plot(discipline.years.plot)
dev.off()

targets <- c("PROPORTIONAL_ATTENDANCE_K_8_PCT","PROPORTIONAL_ATTENDANCE_IEP_PCT")

discipline.attendance.long.dt <- melt(
  discipline.map.dt,
  measure.vars = targets,
  variable.name = "attendance_type",
  value.name = "attendance_rate"
  )

attendance.change.plot <- ggplot() +
  geom_boxplot(
    data = discipline.attendance.long.dt,
    mapping = aes(x = n, y = attendance_rate, fill = attendance_type)
    ) +
  facet_grid(currentSchoolYear ~ .) +
  xlab("Years in Program in year, n = number of districts in 2021") +
  ylab("") +
  labs(title = "Attendance boxplot by years in DCI program")

filename <- paste(dest, "attendance_by_year.png", sep = "")
png(filename = filename, width = 10, height = 8, unit = "in", res = 200)
plot(attendance.change.plot)
dev.off()
```

As all of the previous comparisons show that the schools in the DCI programs have no significant improvement over those outside of it, and in many cases are worse, we inspect whether these schools were also below performing worse before the DCI programs were implemented. 
```{r}
dci.years.dt[, total.years.2022 := max(.SD$year.in.program), by = State.District.ID]

dci.total.years <- unique(dci.years.dt[, .(State.District.ID, total.years.2022)])

comparison.map.dt <- district.map.dt[CONTENT_AREA == "Eng. Language Arts" & GRADE_LEVEL == "03"]

comparison.map.dt <- comparison.map.dt[dci.total.years, on = c("State.District.ID"), nomatch = NULL]


state.average.dt <- fread("./Data Sources CSV/MAP State Average.csv")
state.average.dt[Type == "IEP (exclude MAPA)", Type := "IEP Non MAPA"]

state.average.dt[, PROF_AND_ADV_PCT_STATE := ((Advanced + Proficient) / Reportable) * 100]

comparison.map.dt <- comparison.map.dt[state.average.dt, on = c("YEAR" = "Year", "TYPE" = "Type"), nomatch = NULL]

comparison.map.dt[, PROFICIENT_PCT := as.numeric(PROFICIENT_PCT)]
comparison.map.dt[, ADVANCED_PCT := as.numeric(ADVANCED_PCT)]
comparison.map.dt[, YEAR := as.factor(YEAR)]

comparison.map.dt[, PROF_AND_ADV_PCT := PROFICIENT_PCT + ADVANCED_PCT]
comparison.map.dt[, state.avg.diff := PROF_AND_ADV_PCT - PROF_AND_ADV_PCT_STATE]


yearly.map.diff.by.cohort.plot <-  ggplot(comparison.map.dt) +
  geom_boxplot(
    mapping = aes(x = YEAR, y = state.avg.diff, fill = TYPE)
  ) +
    facet_grid(
      total.years.2022 ~ .
    )

filename <- paste(dest, "map_by_year_and_cohort.png", sep = "")
png(filename = filename, width = 6, height = 12, unit = "in", res = 200)
plot(yearly.map.diff.by.cohort.plot)
dev.off()
```